---
title: "Multi Amplicon - a small example"
author: "Emanuel Heitlinger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Multi Amplicon - a small example

## Preparing sequencing read files and a primer set

The MultiAmplicon package used dedicated classes to ensure that input
to it's `sortAmplicons` function is provided correctly. This means for
the user that file names (and complete paths the directory they are
stored in) of forward and reverse read files should mussed be provided
as a `PairedReadFileSet`. Similarly, primer pairs must be provided in
a `PrimerPairsSet`. We first have a look at how these are generated.

```{r buildup}
library(MultiAmplicon)

primerF <- c(Amp1F = "AGAGTTTGATCCTGGCTCAG", Amp2F = "ACTCCTACGGGAGGCAGC",
             Amp3F = "GAATTGACGGAAGGGCACC", Amp4F = "YGGTGRTGCATGGCCGYT",
             Amp5F = "AAAAACCCCGGGGGGTTTTT", Amp6F = "AGAGTTTGATCCTGCCTCAG")

primerR <- c(Amp1R = "CTGCWGCCNCCCGTAGG", Amp2R = "GACTACHVGGGTATCTAATCC",
             Amp3R = "AAGGGCATCACAGACCTGTTAT", Amp4R = "TCCTTCTGCAGGTTCACCTAC",
             Amp5R = "AAAAACCCCGGGGGGTTTTT", Amp6R = "CCTACGGGTGGCAGATGCAG")

PPS <- PrimerPairsSet(primerF, primerR)

fastq.dir <- system.file("extdata", "fastq", package = "MultiAmplicon")
fastq.files <- list.files(fastq.dir, full.names=TRUE)

Ffastq.file <- fastq.files[grepl("F_filt", fastq.files)]
Rfastq.file <- fastq.files[grepl("R_filt", fastq.files)]

PRF <- PairedReadFileSet(Ffastq.file, Rfastq.file)

MA <- MultiAmplicon(PPS, PRF)

```

## Sorting into multiple amplicons 

Based on the primer sequences we can now sort the sequences into
amplicons. This is done by the function `sortAmplicons`. 


```{r plot, fig.show='hold'}
plot(1:10)
plot(10:1)
```

## Preparing sequencing read files and a primer set

Now wer are 


You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$,
footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

